---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Creating `litr`

<!-- badges: start -->
<!-- badges: end -->

**Note:** If you are looking to use the `litr` package, please visit the `litr` website [here](https://jacobbien.github.io/litr-project/), where you'll be able to learn about the package. If you are interested in how `litr` is made, you are in the right place.

## Background

The `litr` R package allows one to write R packages using literate programming.  The developers of `litr` are believers in literate programming and so, quite naturally, want to use literate programming to develop `litr`.  For obvious reasons, we couldn't use `litr` before it existed.  So we wrote the initial viable version in a standard, non-literate way and released this as [v0.0.1](https://github.com/jacobbien/litr-project/releases/tag/v0.0.1).  We can think of this as the "base case."  This is the last version that is written in the traditional way.  Every subsequent version of `litr` will be generated using the previous version of `litr`.  In particular, v0.0.2 of `litr` is functionally equivalent to v0.0.1, but the package is defined in a generating .Rmd file; a call to v0.0.1's `litr::render()` outputs the `litr` package v0.0.2.  This all might sound complicated, but one can think of this as how developers of an operating system probably write their code on a computer that is running the previous stable version of the operating system.

## How to generate a new version of `litr` using the previous release

To create a new version of `litr`, we first install the latest release, then make any desired changes to `create-litr.Rmd` and then use the installed version's `litr::render()` to create the new version:

```{r}
remotes::install_github("jacobbien/litr-project@*release", subdir = "litr")
litr::render("create-litr.Rmd")
```

In the above code, `@*release` stands for the latest release.  For example, at the time of creating version `0.0.2`, this would be `v0.0.1`.

This will generate the new version of `litr`.  From there, you can build/install as you would for any other package.  For checking the package, use 

```{r}
devtools::check("litr", document = FALSE)
```

The `document = FALSE` prevents `devtools` from running its version of `document()` internally, which would overwrite the modifications that `litr::document()` has made.

For more notes on contributing to `litr`, please see [CONTRIBUTING.md](CONTRIBUTING.md).