---
title: "R Package Templates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{R Package Templates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Predefined `litr` templates make it easy to get started writing R packages of various kinds.  The following table shows the `litr::draft*` functions that can be used to create a `litr` .Rmd file from template.  Click on the "Generated html" and "Generated R package" cells to see what each template produces.

```{r, eval=TRUE, include=FALSE}
library(dplyr)
library(stringr)
```

```{r, eval=TRUE, echo=FALSE, results='asis'}
path <- "https://github.com/jacobbien/litr-project/blob/main/examples"
tbl <- tibble(
  template = c("make-an-r-package", 
               "make-an-r-package-with-data",
               "make-an-r-package-with-rcpp",
               "make-an-r-package-with-extras",
               "make-an-r-package-from-bookdown",
               "make-an-r-package-with-armadillo"),
  pkg_name = c("rhello", "rhasdata", "withrcpp", "withpkgdown", "frombookdown",
               "witharmadillo"),
  description = c("the most basic package: a function and a test",
                  "a package with a data set in it",
                  "a package using `Rcpp`",
                  "a package with a README, vignette(s), a pkgdown site, and a hex sticker",
                  "a package defined in a bookdown",
                  "a package using `RcppArmadillo`"),
  draft = c("`draft()`",
            "`draft_data()`",
            "`draft_rcpp()`",
            "`draft_extras()`",
            "`draft_bookdown()`",
            "`draft_armadillo()`")) %>% 
  transmute(
    `Function` = draft,
    `Template name` = template,
    Description = description,
    `Rmd source file` = str_glue("[create-{pkg_name}.Rmd]({path}/{template}/create-{pkg_name}.Rmd)"),
    `Generated html` = str_glue(
      "[create-{pkg_name}.html](https://htmlpreview.github.io/?{path}/{template}/create-{pkg_name}.html)"
      ),
    `Generated R package` = str_glue("[{pkg_name}/]({path}/{template}/{pkg_name})")
  )
ibookdown <- str_which(tbl$`Template name`, "make-an-r-package-from-bookdown")
tbl[ibookdown, "Rmd source file"] <- str_glue(
  "[index.Rmd]({file_index}), [1description.Rmd]({file_desc}), ..., [4end.Rmd]({file_end})",
  file_index = paste0(path, "/make-an-r-package-from-bookdown/index.Rmd"),
  file_desc = paste0(path, "/make-an-r-package-from-bookdown/1description.Rmd"),
  file_end = paste0(path, "/make-an-r-package-from-bookdown/4end.Rmd"))
tbl[ibookdown, "Generated html"] <- str_glue(
  "[_book/index.html]({file})",
  file = paste0("https://htmlpreview.github.io/?", path, "/make-an-r-package-from-bookdown/_book/index.html"))
tbl[ibookdown, "Generated R package"] <- str_glue(
  "[frombookdown/]({file})",
  file = paste0(path, "/make-an-r-package-from-bookdown/frombookdown"))
knitr::kable(tbl %>% select(-`Template name`))
```

Once you've chosen a template, you can get started by calling the corresponding `litr::draft*` function from the table above.

For example, for the most basic template, you would do the following:

```{r, eval=FALSE}
litr::draft("rhello")
litr::render("create-rhello.Rmd")
```

And to make an R package that uses `Rcpp`, you would start with the following:

```{r, eval=FALSE}
litr::draft_rcpp("withrcpp")
litr::render("create-withrcpp.Rmd")
```
