---
title: "Creating the ``r params$package_name`` R package"
author: "Your Name"
date: "The Date"
knit: litr::render
params:
  package_name: "rhello" # <-- change this to your package name
  package_parent_dir: "." # <-- relative to this file's location
---

<!-- This Rmd file contains all the code needed to define an R package.  Press "Knit" in RStudio or more generally run `rmarkdown::render("name-of-this-file.Rmd")` to generate the R package.  Remember that when you want to modify anything about the R package, you should modify this document rather than the package that is outputted.
-->

```{r knitr-setup, include=FALSE}
# Leave this chunk as is.  It creates a directory for the R package, which will 
# be the base directory for the rest of this file, and then sets things up so
# that code from this document ends up in the R package:
package_name <- params$package_name
package_dir <- file.path(params$package_parent_dir, package_name)
setup_chunk_label <- knitr::opts_current$get("label")
litr::setup(package_dir)
```


## Package setup

We start by specifying the information needed in the DESCRIPTION file of the R package.

```{r package-setup, message=FALSE, results='hide'}
usethis::create_package(
  path = ".",
  fields = list(
    Package = package_name,
    Version = "0.0.0.9000",
    Title = "A Package That Says Hello",
    Description = "This package says hello.  But it's actual purpose is to show how an R package can be completely coded in a single R markdown file.",
    `Authors@R` = person(
      given = "First",
      family = "Last",
      email = "you@gmail.com",
      role = c("aut", "cre")
      )
  )
)
usethis::use_mit_license(copyright_holder = "F. Last")
usethis::use_testthat()
```

## Now to the package itself

### Define a function

Let's define a function for our R package:

```{r}
#' Say hello to someone
#' 
#' @param name Name of a person
#' @param exclamation Whether to include an exclamation mark
#' @export 
say_hello <- function(name, exclamation = TRUE) {
  paste0("Hello ", name, ifelse(exclamation, "!", "."))
}
```

Code chunks whose first line starts with `#'` are added to the package.

We can try running it.

```{r}
say_hello("Jacob")
```

That code chunk does not start with `#'`, so it is not added to the package.

Let's write some tests to make sure the function behaves as desired:

```{r}
testthat::test_that("say_hello works", {
  testthat::expect_equal(say_hello("Jacob"), "Hello Jacob!")
  testthat::expect_equal(say_hello("Jacob", exclamation = FALSE), "Hello Jacob.")
})
```

Code chunks that have `testthat::` in them are added to the package as tests.

## Documenting the package and building

We finish by running commands that will document, build, and install the package.  It may also be a good idea to check the package from within this file.

```{r}
rm(list = ls())
litr::document() # <-- use instead of devtools::document()
devtools::build()
devtools::install()
# devtools::check()
```

