---
title: "Creating the ``r params$package_name`` R package"
author: "Your Name"
date: "The Date"
knit: litr::render
output: litr::litr_html_document
params:
  package_name: "witharmadillo" # <-- change this to your package name
  package_parent_dir: "." # <-- relative to this file's location
---

<!-- This Rmd file contains all the code needed to define an R package.  Press "Knit" in RStudio or more generally run `litr::render("name-of-this-file.Rmd")` to generate the R package.  Remember that when you want to modify anything about the R package, you should modify this document rather than the package that is outputted.
-->

## Package setup

We start by specifying the information needed in the DESCRIPTION file of the R package.

```{r package-setup, message=FALSE, results='hide'}
usethis::create_package(
  path = ".",
  fields = list(
    Package = params$package_name,
    Version = "0.0.0.9000",
    Title = "A Package That Uses RcppArmadillo Code",
    Description = "This package uses RcppArmadillo.",
    `Authors@R` = person(
      given = "First",
      family = "Last",
      email = "you@gmail.com",
      role = c("aut", "cre")
      )
  )
)
usethis::use_mit_license(copyright_holder = "F. Last")
```


## Now to the package itself

To add a C++ function to our R package that uses `RcppArmadillo`, it's the same idea as for regular `Rcpp` (see the `withrcpp` template for more).  The only difference is that the first three lines of the code chunk are specific to `RcppArmadillo`.  Here's an example:

```{Rcpp}
// [[Rcpp::depends(RcppArmadillo)]]
#include <RcppArmadillo.h>
using namespace arma;

//' Cholesky decomposition
//' 
//' @param X A positive definite matrix
//' @export
// [[Rcpp::export]]
arma::mat my_chol(arma::mat& X) {
  return arma::chol(X);
}
```

After at least one `Rcpp` code chunk,^[If we had included a `package_doc` code chunk, then the location of this command would not actually matter.] we next call the following `usethis` command:

```{r}
usethis::use_rcpp_armadillo()
```

And that's it.  If we like, we may call the function that we defined within the Rmarkdown file:

```{r}
x <- matrix(0, 3, 3)
x[upper.tri(x, diag = TRUE)] <- 1:6
my_chol(crossprod(x))
```

However, if we don't need to call the function within the Rmd file, we can set `eval=FALSE`.  This will speed up the knitting process a bit (and the code will still be included in the package).

Let's write some tests to make sure the function behaves as desired:

```{r}
testthat::test_that("my_chol works", {
  x <- matrix(0, 3, 3)
  x[upper.tri(x, diag = TRUE)] <- 1:6
  xchol <- my_chol(crossprod(x))
  testthat::expect_equal(crossprod(xchol), crossprod(x))
  testthat::expect_true(all(xchol[lower.tri(xchol)] == 0))
})
```

Recall that code chunks that have one or more lines starting with `test_that(` (or `testthat::test_that(`) are added to the package as tests.

## Documenting the package and building

We finish by running commands that will document, build, and install the package.  It may also be a good idea to check the package from within this file.

```{r, results='hide'}
litr::document() # <-- use instead of devtools::document()
```


