---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# litr: Writing R packages via literate programming

<!-- badges: start -->
<!-- badges: end -->

The goal of `litr` is to allow you to write a complete R package in a single R markdown document. This enables a workflow for writing R packages that is probably very different from what you are used to.

## Why would one want to write R packages in this way?

When you try to understand the code in an R package, the logic of how the functions relate to each other is often not obvious.  While including function documentation, vignettes, and unit tests are all best practices, they do not convey the chain of logic in the mind of the programmer that went into writing the different functions.  It can be difficult to know how to look through all the functions within even a well-documented R package.  The fact that the functions appear in different files and that functions within files can be defined in arbitrary order makes it unclear how to approach reading the code.  Furthermore, tests are stored in a different place from the functions themselves, making the tests harder to read.  This would all be resolved if we could have a single document that goes through all code and tests in a linear, logical fashion.  Rather than try to construct such a document after the fact, the idea of `litr` is to *make this document the actual source code for the package*.  **The R package is created through the act of knitting this document.**  If we want to modify anything in the package, then we do so by modifying this document and re-knitting.

The above motivation is that of [literate programming](https://en.wikipedia.org/wiki/Literate_programming), introduced by [Donald Knuth](https://www-cs-faculty.stanford.edu/~knuth/), and the direct inspiration is [fast.ai](https://www.fast.ai/about/)'s [`nbdev`](https://nbdev.fast.ai/), which is available in Python.  The `litr` package relies heavily on a number of great tools in R, especially [`knitr`](https://yihui.org/knitr/), [`rmarkdown`](https://rmarkdown.rstudio.com/docs/index.html), [`usethis`](https://usethis.r-lib.org/), [`devtools`](https://devtools.r-lib.org/), and [`testthat`](https://testthat.r-lib.org/).  I should also note that [Yihui Xie](https://yihui.org/en/) has [a post](https://yihui.org/rlp/) where he demonstrates a similar idea, although his approach appears to be more of a proof of concept.

## Getting started

You can install `litr` and get started like so:

```{r}
remotes::install_github("jacobbien/litr-project", subdir = "litr")
rmarkdown::draft("create-rhello.Rmd", template = "make-an-r-package", package = "litr")
```

This installs `litr` and then creates an R markdown template file called `create-rhello.Rmd` that demonstrates the literate programming workflow for writing an R package.  In particular, when you knit `create-rhello.Rmd`, it creates a tiny example R package called `rhello` with one function and one test function.  To knit, you can either press "Knit" in RStudio or use the following command:

```{r}
litr::render("create-rhello.Rmd")
```

## Examples: Package templates

Predefined `litr` templates make it easy to get started writing R packages of various kinds.  The following table shows the package templates available.  Click on the "Generated html" and "Generated R package" cells to see what each template produces.

```{r, eval=TRUE, include=FALSE}
library(dplyr)
library(stringr)
```

```{r, eval=TRUE, echo=FALSE, results='asis'}
path <- "https://github.com/jacobbien/litr-project/blob/main/examples"
tibble(
  template = c("make-an-r-package", "make-an-r-package-with-data", "make-an-r-package-with-rcpp"),
  pkg_name = c("rhello", "rhasdata", "withrcpp"),
  description = c("the most basic package: a function and a test",
                  "a package with a data set in it",
                  "a package using `Rcpp`")) %>% 
  transmute(
    `Template name` = template,
    Description = description,
    `Rmd source file` = str_glue("[create-{pkg_name}.Rmd]({path}/{template}/create-{pkg_name}.Rmd)"),
    `Generated html` = str_glue(
      "[create-{pkg_name}.html](https://htmlpreview.github.io/?{path}/{template}/create-{pkg_name}.html)"
      ),
    `Generated R package` = str_glue("[{pkg_name}/]({path}/{template}/{pkg_name})")
  ) %>% 
knitr::kable()
```

Once you've chosen a template, you can get started as follows:

```{r}
rmarkdown::draft("create-[name-of-your-package].Rmd", template = "[name-of-template]", package = "litr")
litr::render("create-[name-of-your-package].Rmd")
```

