---
title: "A Basic Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A Basic Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

\newcommand{\real}{\mathbb R}

We walk through the basics of `litr` with a simple example.

Suppose we want to create a package called `rleastsquares` that does least squares.  To do so, we will create a .Rmd file called something like `create-rleastsquares.Rmd`.  We recommend creating this from a template: 

```{r, eval=FALSE}
rmarkdown::draft("create-rleastsquares.Rmd",
                 template = "make-an-r-package",
                 package = "litr")
```

When you work from template, you'll notice some special lines in the yaml:

```
knit: litr::render
output: litr::litr_html_document
params:
  package_name: "rhello" # <-- change this to your package name
  package_parent_dir: "." # <-- relative to this file's location
```

This is what will lead to an R package being created when you press "Knit".  We'll replace `"rhello"` with  `"rleastsquares"`.

## Package setup

> **Note:** Every R package needs a DESCRIPTION file.  We'll start by filling in the relevant information.

```{r package-setup, message=FALSE, results='hide', eval=FALSE}
usethis::create_package(
  path = ".",
  fields = list(
    Package = param$package_name,
    Version = "0.0.0.9000",
    Title = "Fit Least Squares",
    Description = "A package that fits least squares.",
    `Authors@R` = person(
      given = "First",
      family = "Last",
      email = "you@gmail.com",
      role = c("aut", "cre")
      )
  )
)
usethis::use_mit_license(copyright_holder = "F. Last")
```

## Writing a function for the package

Since this is an R markdown file, we can use latex to explain our code, provide derivations, etc.

Suppose we have a response vector $y\in\real^n$ and a data matrix $X\in\real^{n\times p}$.

We want to find the solution to the problem

$$
\min_{\beta\in\real^p}\|y-X\beta\|^2
$$

We'll assume that $X$ is full rank with $n > p$.  We know that the solution is given by

$$
\hat\beta=(X^TX)^{-1}X^Ty.
$$

We'll write a function that does exactly that!

```{r}
#' Get the OLS solution
#' 
#' @param y our response, which is an n-vector
#' @param X our data matrix, which is n by p
#' @export
do_least_squares <- function(y, X) {
  if(nrow(X) != length(y))
    stop("The number of rows of X must match the length of y.")
  as.numeric(solve(crossprod(X), crossprod(X, y)))
}
```

> **Note:** Code chunks whose first line starts with `#'` are added to the package.  If you're not familiar with `roxygen2`, see [here](https://roxygen2.r-lib.org/) for more.

Now that we've defined `do_least_squares()`, let's try it out!

```{r}
set.seed(123)
n <- 100
p <- 1
x <- cbind(1, matrix(rnorm(n*p), n, p))
beta_star <- c(2, 0.5)
sigma <- 0.1
y <- x %*% beta_star + sigma * rnorm(n)
```

> **Note:** This code chunk does not start with `#'`, so it is not added to the package.

```{r}
betahat <- do_least_squares(y, x)
```

```{r}
plot(x[, 2], y)
abline(betahat[1], betahat[2], col = 2, lwd=2)
```

Let's see how this compares to `lm`'s answer.


```{r}
fit_lm <- lm(y ~ x[, 2])
```

```{r}
fit_lm$coefficients
```

Compare that to... 

```{r}
betahat
```

Ok, `do_least_squares()` appears to be working.  Let's define a formal unit test based on the example above.

```{r}
testthat::test_that("do_least_squares() works", {
  set.seed(123)
  n <- 100
  p <- 1
  x <- cbind(1, matrix(rnorm(n*p), n, p))
  beta_star <- c(2, 0.5)
  sigma <- 0.1
  y <- x %*% beta_star + sigma * rnorm(n)
  fit_lm <- lm(y ~ x[, 2])
  # do lm and our function give the same coefficient vector?
  testthat::expect_equal(do_least_squares(y, x),
                         as.numeric(fit_lm$coefficients))
  # do we get the desired error when there is a length mismatch?
  testthat::expect_error(do_least_squares(y[-1], x), "must match")
})
```

> **Note:** Code chunks that have one or more lines starting with `test_that(` or `testthat::test_that(` are added to the package as tests.

## Some fancier features

### Finer control over where in the package your code is sent

As noted above, `litr` detects whether to send a code chunk to the package based on whether it starts with `#'` or has `test_that` in it.  However, sometimes you'll want finer control.  In this case you can override the behavior by explicitly specifying the target location.  To do so, use a code chunk option of the form `send_to="R/myfile.R"`.  This will add your code to that particular file (either creating it if need be or else appending it).  There are two primary use cases for this feature: (a) when you don't want to use `roxygen2` to document a function and (b) when you want several functions to appear together in the same .R file.

### Using a function from a different package

Imagine we wanted to actually use a function from another package in our own.  For example, perhaps we want to use `lsfit()` from the `stats` package:

```{r}
#' Get the OLS solution using lsfit()
#' 
#' @param y our response, which is an n-vector
#' @param X our data matrix, which is n by p
#' @export
do_least_squares_with_lsfit <- function(y, X) {
  fit <- stats::lsfit(x = X, y = y, intercept = FALSE)
  return(as.numeric(fit$coefficients))
}
```

And then we also update the DESCRIPTION file with this package dependence:

```{r, eval=FALSE}
usethis::use_package("stats")
```

> To use a function from another package, simply use the `pkg::` prefix when calling it and then add `usethis::use_package("pkg")` to include the package dependency.

### Including a dataset in your package

There's a [template](package-templates.html) for that.

### Using `Rcpp` in your package

There's a [template](package-templates.html) for that.

### Including a README, vignettes, and `pkgdown` site

There's a [template](package-templates.html) for that.

### Defining your package with `bookdown`

If you're writing a large package, it may be convenient to define it across multiple .Rmd files. You can use `bookdown` for this, which leads to a nice looking online book with multiple chapters.  There's a [template](package-templates.html) for that.

## Documenting the package

We finish by running a commands that will turn the `royxgen2` into traditional documentation files in our R package.

```{r, eval=FALSE}
litr::document() # <-- use instead of devtools::document()
```

